---
title: "Extended Analysis"
author: "Matt McKnight"
date: "June 2, 2017"
output:
  html_document:
    df_print: paged
    keep_md: true
    theme: cerulean
    highlight: haddock
    toc: yes
    toc_float: yes
    code_fold: hide
  pdf_document: default
---

# NONPROFIT SERVICE AREAS IN TOP 5 CITIES

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F, fig.width=10)

library( ggmap )
library( rgdal )
library( devtools )
library( data.table )
library( censusapi )
library( plyr )
library( gdata )
library( geojsonio )
library( acs )
library( tigris )
library( sf ) 
library( sp )  
library( rgeos )
library( spatialEco ) 
library( leaflet )
library( raster )
library( repmis )
library( TSP )
library( magrittr )
library( lattice )
library( GISTools )

```


## Download
Download master file of tax exempt organizations from the web using the fread function in the data.table. This function can take data directly from the web and read it directly into R as a data frame. This step takes a while to download because the file is so large (over 1.5 million organizations).

```{r eval=FALSE}

print("Don't run me")

# Download master file of tax exempt organizations
orgs <- fread("http://nccs-data.urban.org/data/bmf/2016/bmf.bm1608.csv")

dim(orgs)
head(orgs)

save(orgs, file = "../ASSETS/orgs.Rda")
```

## Subset by MSA

```{r eval=FALSE}

print("Don't run me")


# Top five MSAs by population

# Location for all FIPS codes and their respective MSAs
# https://www.bea.gov/regional/docs/msalist.cfm#N

# 1) New York-Newark-Jersey City, NY-NJ-PA

nyc_orgs <- subset(orgs, 
                     grepl("34003", FIPS) | 
                     grepl("34013", FIPS) |
                     grepl("34017", FIPS) |
                     grepl("34019", FIPS) |
                     grepl("34023", FIPS) |
                     grepl("34025", FIPS) |
                     grepl("34027", FIPS) |
                     grepl("34029", FIPS) |
                     grepl("34031", FIPS) |
                     grepl("34035", FIPS) |
                     grepl("34037", FIPS) |
                     grepl("34039", FIPS) |
                     grepl("36005", FIPS) |
                     grepl("36027", FIPS) |
                     grepl("36047", FIPS) |
                     grepl("36059", FIPS) |
                     grepl("36061", FIPS) |
                     grepl("36071", FIPS) |
                     grepl("36079", FIPS) |
                     grepl("36081", FIPS) |
                     grepl("36085", FIPS) |
                     grepl("36087", FIPS) |
                     grepl("36103", FIPS) |
                     grepl("36119", FIPS) |
                     grepl("42103", FIPS) )

# 2) Los Angeles-Long Beach-Anaheim, CA 

la_orgs <- subset(orgs, 
                     grepl("6037", FIPS) | 
                     grepl("6059", FIPS) )

# 3) Chicago-Naperville-Elgin, IL-IN-WI 

chi_orgs <- subset(orgs, 
                     grepl("17031", FIPS) | 
                     grepl("17037", FIPS) |
                     grepl("17043", FIPS) |
                     grepl("17063", FIPS) |
                     grepl("17089", FIPS) |
                     grepl("17093", FIPS) |
                     grepl("17097", FIPS) |
                     grepl("17111", FIPS) |
                     grepl("17197", FIPS) |
                     grepl("18073", FIPS) |
                     grepl("18089", FIPS) |
                     grepl("18111", FIPS) |
                     grepl("18127", FIPS) |
                     grepl("55059", FIPS) )

# 4) Dallas-Fort Worth-Arlington, TX

dfw_orgs <- subset(orgs, 
                     grepl("48085", FIPS) | 
                     grepl("48113", FIPS) |
                     grepl("48121", FIPS) |
                     grepl("48139", FIPS) |
                     grepl("48221", FIPS) |
                     grepl("48231", FIPS) |
                     grepl("48251", FIPS) |
                     grepl("48257", FIPS) |
                     grepl("48367", FIPS) |
                     grepl("48397", FIPS) |
                     grepl("48425", FIPS) |
                     grepl("48439", FIPS) |
                     grepl("48497", FIPS) )
                     
# 5) Houston-The Woodlands-Sugar Land, TX 

hou_orgs <- subset(orgs, 
                     grepl("48015", FIPS) | 
                     grepl("48039", FIPS) |
                     grepl("48071", FIPS) |
                     grepl("48157", FIPS) |
                     grepl("48167", FIPS) |
                     grepl("48201", FIPS) |
                     grepl("48291", FIPS) |
                     grepl("48339", FIPS) |
                     grepl("48473", FIPS) )
                     

```

## Subset by poverty organizations

```{r eval=FALSE}

print("Don't run me")

# 1) New York-Newark-Jersey City, NY-NJ-PA

nyc_pov_orgs <- subset(nyc_orgs, 
                        grepl("B6.", NTEECC) | 
                        grepl("B92", NTEECC) |
                        grepl("E32", NTEECC) |  
                        grepl("E7.", NTEECC) |  
                        grepl("F2.", NTEECC) |
                        grepl("F32", NTEECC) |  
                        grepl("I2.", NTEECC) |
                        grepl("I4.", NTEECC) |  
                        grepl("I7.", NTEECC) |
                        grepl("I8.", NTEECC) |  
                        grepl("J2.", NTEECC) |
                        grepl("J3.", NTEECC) |  
                        grepl("K3.", NTEECC) |  
                        grepl("K4.", NTEECC) |  
                        grepl("K5.", NTEECC) |
                        grepl("L2.", NTEECC) |
                        grepl("L3.", NTEECC) |  
                        grepl("L4.", NTEECC) |  
                        grepl("L8.", NTEECC) |  
                        grepl("O2.", NTEECC) |  
                        grepl("O3.", NTEECC) |
                        grepl("O5.", NTEECC) |  
                        grepl("P", NTEECC)   |
                        grepl("R2.", NTEECC) |  
                        grepl("R3.", NTEECC) |
                        grepl("S2.", NTEECC) )

dim(nyc_pov_orgs)
head(nyc_pov_orgs)

save(nyc_pov_orgs, file = "../ASSETS/nyc_pov_orgs.Rda")

# 2) Los Angeles-Long Beach-Anaheim, CA 

la_pov_orgs <- subset(la_orgs, 
                        grepl("B6.", NTEECC) | 
                        grepl("B92", NTEECC) |
                        grepl("E32", NTEECC) |  
                        grepl("E7.", NTEECC) |  
                        grepl("F2.", NTEECC) |
                        grepl("F32", NTEECC) |  
                        grepl("I2.", NTEECC) |
                        grepl("I4.", NTEECC) |  
                        grepl("I7.", NTEECC) |
                        grepl("I8.", NTEECC) |  
                        grepl("J2.", NTEECC) |
                        grepl("J3.", NTEECC) |  
                        grepl("K3.", NTEECC) |  
                        grepl("K4.", NTEECC) |  
                        grepl("K5.", NTEECC) |
                        grepl("L2.", NTEECC) |
                        grepl("L3.", NTEECC) |  
                        grepl("L4.", NTEECC) |  
                        grepl("L8.", NTEECC) |  
                        grepl("O2.", NTEECC) |  
                        grepl("O3.", NTEECC) |
                        grepl("O5.", NTEECC) |  
                        grepl("P", NTEECC)   |
                        grepl("R2.", NTEECC) |  
                        grepl("R3.", NTEECC) |
                        grepl("S2.", NTEECC) )

dim(la_pov_orgs)
head(la_pov_orgs)

save(la_pov_orgs, file="../ASSETS/la_pov_orgs.Rda")

# 3) Chicago-Naperville-Elgin, IL-IN-WI 


chi_pov_orgs <- subset(chi_orgs, 
                        grepl("B6.", NTEECC) | 
                        grepl("B92", NTEECC) |
                        grepl("E32", NTEECC) |  
                        grepl("E7.", NTEECC) |  
                        grepl("F2.", NTEECC) |
                        grepl("F32", NTEECC) |  
                        grepl("I2.", NTEECC) |
                        grepl("I4.", NTEECC) |  
                        grepl("I7.", NTEECC) |
                        grepl("I8.", NTEECC) |  
                        grepl("J2.", NTEECC) |
                        grepl("J3.", NTEECC) |  
                        grepl("K3.", NTEECC) |  
                        grepl("K4.", NTEECC) |  
                        grepl("K5.", NTEECC) |
                        grepl("L2.", NTEECC) |
                        grepl("L3.", NTEECC) |  
                        grepl("L4.", NTEECC) |  
                        grepl("L8.", NTEECC) |  
                        grepl("O2.", NTEECC) |  
                        grepl("O3.", NTEECC) |
                        grepl("O5.", NTEECC) |  
                        grepl("P", NTEECC)   |
                        grepl("R2.", NTEECC) |  
                        grepl("R3.", NTEECC) |
                        grepl("S2.", NTEECC) )

dim(chi_pov_orgs)
head(chi_pov_orgs)

save(chi_pov_orgs, file="../ASSETS/chi_pov_orgs.Rda")

# 4) Dallas-Fort Worth-Arlington, TX

dfw_pov_orgs <- subset(dfw_orgs, 
                        grepl("B6.", NTEECC) | 
                        grepl("B92", NTEECC) |
                        grepl("E32", NTEECC) |  
                        grepl("E7.", NTEECC) |  
                        grepl("F2.", NTEECC) |
                        grepl("F32", NTEECC) |  
                        grepl("I2.", NTEECC) |
                        grepl("I4.", NTEECC) |  
                        grepl("I7.", NTEECC) |
                        grepl("I8.", NTEECC) |  
                        grepl("J2.", NTEECC) |
                        grepl("J3.", NTEECC) |  
                        grepl("K3.", NTEECC) |  
                        grepl("K4.", NTEECC) |  
                        grepl("K5.", NTEECC) |
                        grepl("L2.", NTEECC) |
                        grepl("L3.", NTEECC) |  
                        grepl("L4.", NTEECC) |  
                        grepl("L8.", NTEECC) |  
                        grepl("O2.", NTEECC) |  
                        grepl("O3.", NTEECC) |
                        grepl("O5.", NTEECC) |  
                        grepl("P", NTEECC)   |
                        grepl("R2.", NTEECC) |  
                        grepl("R3.", NTEECC) |
                        grepl("S2.", NTEECC) )

dim(dfw_pov_orgs)
head(dfw_pov_orgs)

save(dfw_pov_orgs, file="../ASSETS/dfw_pov_orgs.Rda")

# 5) Houston-The Woodlands-Sugar Land, TX 

hou_pov_orgs <- subset(hou_orgs, 
                        grepl("B6.", NTEECC) | 
                        grepl("B92", NTEECC) |
                        grepl("E32", NTEECC) |  
                        grepl("E7.", NTEECC) |  
                        grepl("F2.", NTEECC) |
                        grepl("F32", NTEECC) |  
                        grepl("I2.", NTEECC) |
                        grepl("I4.", NTEECC) |  
                        grepl("I7.", NTEECC) |
                        grepl("I8.", NTEECC) |  
                        grepl("J2.", NTEECC) |
                        grepl("J3.", NTEECC) |  
                        grepl("K3.", NTEECC) |  
                        grepl("K4.", NTEECC) |  
                        grepl("K5.", NTEECC) |
                        grepl("L2.", NTEECC) |
                        grepl("L3.", NTEECC) |  
                        grepl("L4.", NTEECC) |  
                        grepl("L8.", NTEECC) |  
                        grepl("O2.", NTEECC) |  
                        grepl("O3.", NTEECC) |
                        grepl("O5.", NTEECC) |  
                        grepl("P", NTEECC)   |
                        grepl("R2.", NTEECC) |  
                        grepl("R3.", NTEECC) |
                        grepl("S2.", NTEECC) )

dim(hou_pov_orgs)
head(hou_pov_orgs)

save(hou_pov_orgs, file="../ASSETS/hou_pov_orgs.Rda")
```

## Geocode addresses of poverty organizations

```{r}

# download specific ggmap, register Google key

devtools::install_github("dkahle/ggmap")

register_google(key = 'AIzaSyAMrECdXqpOt443ms6nzh118uUsqC2lE6M',
                account_type = "premium", day_limit = 15000)

# subset orgs for gecoding address information

nyc_pov_orgs_gps <- subset(nyc_pov_orgs, select=c(EIN:INCOME))

nyc_pov_orgs_gps$whole_address <- paste(nyc_pov_orgs_gps$ADDRESS,                                                        nyc_pov_orgs_gps$CITY, sep = ", ") 
nyc_pov_orgs_gps$whole_address <- paste(nyc_pov_orgs_gps$whole_address,                                                  nyc_pov_orgs_gps$STATE, sep = ", ")
nyc_pov_orgs_gps$whole_address <- paste(nyc_pov_orgs_gps$whole_address,                                                  nyc_pov_orgs_gps$ZIP5, sep = ", ") 

nyc_pov_orgs_gps <- subset(nyc_pov_orgs_gps, select=c(whole_address, EIN:INCOME))

for (i in 1:nrow(nyc_pov_orgs_gps)) {
  latlon = geocode(as.character(nyc_pov_orgs_gps[i,1]), override_limit=TRUE) 
  nyc_pov_orgs_gps$lon[i] = as.numeric(latlon[1])
  nyc_pov_orgs_gps$lat[i] = as.numeric(latlon[2])
}


save(nyc_pov_orgs_gps, file="../ASSETS/nyc_pov_orgs_gps.Rda")

##############################3

# 5) Houston MSA geocoding

register_google(key = 'AIzaSyAMrECdXqpOt443ms6nzh118uUsqC2lE6M',
                account_type = "premium", day_limit = 15000)

# subset orgs for gecoding address information

hou_pov_orgs_gps <- subset(hou_pov_orgs, select=c(EIN:INCOME))

hou_pov_orgs_gps$whole_address <- paste(hou_pov_orgs_gps$ADDRESS,                                                        hou_pov_orgs_gps$CITY, sep = ", ") 
hou_pov_orgs_gps$whole_address <- paste(hou_pov_orgs_gps$whole_address,                                                  hou_pov_orgs_gps$STATE, sep = ", ")
hou_pov_orgs_gps$whole_address <- paste(hou_pov_orgs_gps$whole_address,                                                  hou_pov_orgs_gps$ZIP5, sep = ", ") 

hou_pov_orgs_gps <- subset(hou_pov_orgs_gps, select=c(whole_address, EIN:INCOME))

for (i in 1:nrow(hou_pov_orgs_gps)) {
  latlon = geocode(as.character(hou_pov_orgs_gps[i,1])) 
  hou_pov_orgs_gps$lon[i] = as.numeric(latlon[1])
  hou_pov_orgs_gps$lat[i] = as.numeric(latlon[2])
}

# Make sure you find a way to then take any PO Boxes and place them in the 
# center of centroids, and only the ones that are NAs.

save(hou_pov_orgs_gps, file="../ASSETS/hou_pov_orgs_gps.Rda")
```

# CENSUS DATA

## Census API package

Download US census data for top 5 MSAs

```{r}

censuskey <- "f8064a17832b439e690be95ab0e0ef9a78617584"

var.list <- c("NAME", 
              "B25077_001E", "B25003_001E", "B25003_002E", "B25003_003E",
              "B25002_001E", "B25002_002E", "B25002_003E", "B01003_001E",
              "B19013_001E", "B17020_002E", "B12006_001E", "B23025_005E",                      "B15003_017E", "B15003_018E",
              "B03001_001E", "B03002_003E", "B03002_004E", "B03002_005E",
              "B03002_006E", "B03002_007E", "B03002_008E", "B03002_009E", 
              "B03001_002E", "B03001_003E", 
              "GEOID" 
              )

acs_2015 <- getCensus( name="acs5",
                       vintage = 2015,
                       key = censuskey,
                       vars= var.list, 
                       region ="tract:*", 
                       regionin ="state:36" 
                     )


# 1) subset to NY

acs_2015_syr <- acs_2015[which (acs_2015$county == "067" | 
                                acs_2015$county == "075" |
                                acs_2015$county == "053"), ]

# 2) subset to LA


# 3) subset to CHI


# 4) subset to DFW


# 5) subset to HOU


```


## Rename Variables

Rename variables of interest in each subsetted data

```{r}

rename( acs_2015_syr, 
       c("B25077_001E" = "mdn_hous_val", 
         "B25003_001E" = "tenure_tot", 
         "B25003_002E" = "tenure_own", 
         "B25003_003E" = "tenure_rent",
         "B25002_001E" = "tot_occup", 
         "B25002_002E" = "occupied",
         "B25002_003E" = "vacant", 
         "B01003_001E" = "tot_pop", 
         "B19013_001E" = "mhh_income",
         "B17020_002E" = "poverty",
         "B12006_001E" = "labor_part",
         "B23025_005E" = "unemploy",
         "B15003_017E" = "high_sch",
         "B15003_018E" = "ged",
         "B03001_001E" = "tot_race",
         "B03002_003E" = "white",             
         "B03002_004E" = "black", 
         "B03002_005E" = "am_ind",
         "B03002_006E" = "asian", 
         "B03002_007E" = "islander",
         "B03002_008E" = "other", 
         "B03002_009E" = "mixed",
         "B03001_002E" = "not_hispanic",
         "B03001_003E" = "hispanic")
        )

acs_2015_syr <- rename.vars(acs_2015_syr,
            c("B25077_001E", "B25003_001E", "B25003_002E", "B25003_003E",
              "B25002_001E", "B25002_002E", "B25002_003E", "B01003_001E",
              "B19013_001E", "B17020_002E", "B12006_001E", "B23025_005E",                      "B15003_017E", "B15003_018E",
              "B03001_001E", "B03002_003E", "B03002_004E", "B03002_005E",
              "B03002_006E", "B03002_007E", "B03002_008E", "B03002_009E", 
              "B03001_002E", "B03001_003E"),  
            c("mdn_hous_val", "tenure_tot", "tenure_own", "tenure_rent",
              "tot_occup", "occupied", "vacant", "tot_pop", 
              "mhh_income", "poverty", "labor_part", "unemploy",
              "high_sch", "ged",
              "tot_race", "white", "black", "am_ind",
              "asian", "islander", "other", "mixed",
              "not_hispanic", "hispanic")
            )

head(acs_2015_syr)

```

## Create variables 

Create 8 variables of interest from the US Census data


```{r}

# Create proportion white and proportion minority

# Proportion race variables (by total population per census tract)

# White non-hispanic

acs_2015_syr$porp_white <- (acs_2015_syr$white) / (acs_2015_syr$tot_race)

# All minorities (reverse of white non-hispanic) 

acs_2015_syr$porp_m <- abs((acs_2015_syr$porp_white) - 1) 

# Black

acs_2015_syr$prop_black <- acs_2015_syr$black / acs_2015_syr$tot_race

# Hispanic

acs_2015_syr$prop_hisp <- acs_2015_syr$hispanic / acs_2015_syr$tot_race

# Proportion of people whose incomes are lower than the federal poverty line

acs_2015_syr$pov <- acs_2015_syr$poverty / acs_2015_syr$tot_pop

# percentage of unemployed, NOT umemployment rate

acs_2015_syr$unemp <- acs_2015_syr$unemploy / acs_2015_syr$tot_pop

# high school degree, either graduated or with a GED

acs_2015_syr$educ <- (acs_2015_syr$high_sch + acs_2015_syr$ged) /                                      acs_2015_syr$tot_pop

# Porportion of occupied housing that is rented

acs_2015_syr$prop_rent <- acs_2015_syr$tenure_rent / acs_2015_syr$tot_occup


names(acs_2015_syr)


```

# Create Geojson Files


```{r}

# How to download shapefiles from the census directly using tigris functions
# https://rdrr.io/cran/tigris/man/blocks.html

Mad_county <- tracts(state = "NY", county = "053", year = "2015")

On_county <- tracts(state = "NY", county = "067", year = "2015")

Osw_county <- tracts(state = "NY", county = "075", year = "2015")

# Drop NAs
Mad_county <- na.omit(Mad_county)
On_county <- na.omit(On_county)
Osw_county <- na.omit(Osw_county)

```



## Clean census data GEOID

The data downloaded from the census API includes values in GEOID cells that need to be removed prior to merging with shapefiles.

```{r}

head( acs_2015_syr$GEOID)

acs_2015_syr$GEOID <- gsub( acs_2015_syr$GEOID, pattern="14000US", replacement="" )
census_dat <- acs_2015_syr

head(acs_2015_syr$GEOID)
```


## Merge 3 spatial files per MSA

When the shapefiles are joined, they do not create a full list variable of all GEOIDs. This step creates a GEOID variabe combined from all three shapefiles.

```{r}

# Spatial join
syr_msa <- union(Mad_county, On_county)
syr_msa <- union(syr_msa, Osw_county)

# Create complete 

GEOID <- data.frame(syr_msa$GEOID.1, syr_msa$GEOID.2, syr_msa$GEOID)
head(GEOID)
GEOID_tot <- GEOID[!is.na(GEOID)]
GEOID$GEOID_tot <- GEOID_tot
head(GEOID)

syr_msa$GEOID_full <- GEOID_tot

# syr_msa <- cbind( Mad_county, On_county )
# syr_msa <- cbind( syr_msa, Osw_county )
# Does not work, Error: arguments imply differing number of rows: 16, 140

plot(syr_msa)

```


## Merge census attribute data with shapefile data

```{r}

# Merge data with MSA shp file

syr_merged <- merge(syr_msa, census_dat, 
                    by.x = "GEOID_full", by.y = "GEOID")                      

plot(syr_merged)
title(main = "Syracuse MSA, NY")
head(syr_merged)
nrow(syr_merged)

# Drop census tract 9900, water census tract in Oswego county
syr_merged_clean <- syr_merged[!(syr_merged$tract=="990000"),]
nrow(syr_merged)

plot(syr_merged_clean)
title(main = "Syracuse MSA, NY")

```


# GEOJSON FILES

## Convert shapefiles to geojson files for the Syracuse MSA.

```{r}
geojson_write(syr_merged_clean, geometry="polygon", file="../ASSETS/syr_merged.geojson")

```


## Create a shapefile with centroids for every census tract and convert to geojson format.

```{r}

syr_merged_cen = gCentroid(syr_merged_clean, byid=TRUE)

geojson_write(syr_merged_cen, geometry="point", file="../ASSETS/syr_merged_cen.geojson")

```


### Analysis ###

```{r}

# Create buffers, render in same projection
centroids_buff <- gBuffer(spgeom = centroids, width = .024, joinStyle="ROUND", byid = T)
centroids_buff@proj4string
spTransform(centroids_buff, CRS("+proj=longlat +datum=WGS84"))

org_buff <- gBuffer(spgeom = pov_orgs, width = .024, joinStyle="ROUND", byid = T)
centroids_buff@proj4string
spTransform(org_buff, CRS("+proj=longlat +datum=WGS84"))


# join data to centroid buffers
centroids_buff$ID <- c(1:185)
syr$ID <- c(1:185)
org_buff$ID <- c(1:506)
pov_orgs$ID <- c(1:506)

c_buff <- merge( centroids_buff, syr@data, "ID", "ID" )
test2 <- merge( org_buff, pov_orgs@data, "ID", "ID" )


# Number of nonprofits within a 1.5 miles of middle of census tract.
res <- poly.counts(pov_orgs, c_buff)
setNames(res, c_buff$tract)
syr$output1 <- setNames(res, c_buff$tract)

# Number of nonprofits in each census tract
res2 <- poly.counts(pov_orgs, syr)
setNames(res2, syr$tract)
syr$output2 <- setNames(res2, syr$tract)

# Org way:
# Number of nonprofit service catchment areas within a given census tract

inter_npo <- gIntersects(org_buff, syr, byid=TRUE)
inter_npo2 <- inter_npo + 0
inter_npo3 <- as.data.frame(inter_npo2)
tot_col <- apply(inter_npo3, 1, sum)
syr$output3 <- tot_col

# Scatterplots

plot( syr$porp_white, syr$output3, 
      abline(lm(syr$output3~syr$porp_white)),
      main="Porportion White and # of NPOs per Census Tract",
      xlab="Porportion White", ylab="# of NPOs per Census Tract",
      col="blue", las = 1)

plot( syr$mhh_income, syr$output3, 
      abline(lm(syr$output3~syr$mhh_income)), 
      main="Median HH Income and # of NPOs per Census Tract",
      xlab="Median HH Income", ylab="# of NPOs per Census Tract",
      col="blue", las = 1)

# Regressions with number of NPO service catchment areas by census tract

summary(m1 <- lm(syr$output3 ~ syr$porp_white))
summary(m2 <- lm(syr$output3 ~ syr$pov))
summary(m3 <- lm(syr$output3 ~ syr$educ))
summary(m4 <- lm(syr$output3 ~ syr$prop_rent))
summary(m5 <- lm(syr$output3 ~ syr$mhh_income))

summary(m6 <- lm(syr$output3 ~ syr$porp_white + syr$pov + syr$educ +                              syr$prop_rent + syr$mhh_income))

# poverty as a dependent variable, number of NPOs as a independent variable

summary(m7 <- lm(syr$pov ~ syr$porp_white + syr$output3 + syr$educ +                              syr$prop_rent))

```




